---
title: "Effects of Promotions on the Sale of GM Cereals"
author: "Glen Lewis"
date: "MM/DD/YYYY"   
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include = FALSE}
# Course: 5210 Data Visualization 
# Purpose: Mid Term Project
# Date: 
# Author: Glen Lewis

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Hypothesizes and Assessment Criteria

H1~0~ = Promotions did not impact sales of GM cereal products
<br> H1~A~ = Promotions did impact the sales of GM cereal products

*H1 assess how effective GM promotions are at increasing sales.*

The impact of a type I error would be to create a perceived business environment where decision makers believe their promotion efforts have an influence on increasing sales when they are not.  This environment may lead to decision-making that continues following a strategy that is unknowingly ineffective, leading to flat if not declining sales.  

The impact of a type II error is interesting in that it would great a perceived business environment where decision makers believe their promotion efforts are ineffective at influencing sales when they are.   This environment may lead to the unnecessary adoption of a new strategy or the wasteful allocates of resources to produce an effect that is already occurring.  Those decisions: however, could also expand/grow the positive effects their current promotion strategies are having. 

A type I error could lead to flat sales - a significant opportunity cost – that grows which each passing day. A type II error may or may not have a negative impact on future sales depending; thus, we consider a type I error as the bigger risk.  Therefore, the study will set alpha to .05 minimizing the potential for a type I error when answering H1. 

H2~0~ = GM promotion techniques were not more effective than other Brands' promotion techniques
<br> H2~A~ = GM promotion techniques were more effective than other Brands' promotion techniques.

*H2 assess how effective GM sale promotions are in raising sales vis-a-vis it's competitors' promotions' efforts.*  

The impact of a type I error would create a perceived business environment where decision makers believe their promotion techniques are more effect than the competitions.  This environment may lead to decision-making to continue following a strategy that is ineffective and unknowingly opening market expansion opportunities to competitors. 

The impact of a type II error is like a type II error for H1. While it could lead to the unnecessary allocation of resources, that reallocation could augment the competitive advantage its current efforts are unknowing creating.

A type I error could lead to a loss of market share, something GM could not easily regain without significant effort and cost. A type II error may or may not have a negative impact on future sales depending; thus, we consider a type I error as the bigger risk.  Therefore, the study will set alpha to .05 minimizing the potential for a type I error when answering H2.

# Load and transform data
```{r}
library(tidyverse) # For cleaning data.
library(lubridate) # For date conversions.
library(vtable) # To conduct descriptive EDA.
library(patchwork) # Use to put graphs together (faceting) in the same frame.
library(scales) # To get convert axis scales and labels.
library(here) # For saving plots to specific folders.
library(jtools) # For reviewing and plotting lm relationships.
library(janitor) 
library(ggannotate) # For annotations.
```

# Cleaning and 'Tidying' Data
```{r}
# Following loads the data csv's into the environment via a base_product_data and base_sales_data objects.

base_product_data <- read_csv('mtp_product_data.csv')
base_sales_data <- read_csv('mtp_sales_data.csv')

head(base_product_data, 10)
# NOTE: For small data sets with few variables head easier that vtable. 

vtable(base_sales_data) 
# FOR ACTION: (1) Standardize UPCs in both data sets to act as a key for a join. (2) Join the to data sets into a single base set. (3) Convert the promo and add variables to factors.

wip_product_data <- base_product_data %>% 
  mutate(UPC = str_remove(UPC, '^00-')) %>% 
  # removes the 00- characters from the UPC
  mutate(UPC = str_replace_all(UPC, '-', '\\.')) %>% 
  # replaces the - in the UPC with a '.'.  This is need to setup a key for the join.
  mutate(brand = str_replace(brand, '^GENERAL MILLS', 'GENERAL_MILLS')) %>% 
  # replaces the " " with '_' enabling easy separation and manipulation of variables.
  mutate(brand = str_replace(brand, ' ', '\\\\')) %>% 
  # replaces the first ' ' with a '\' so the col can be easily separated into maker and brand.
  separate(brand, c('maker', 'brand'), sep = '\\\\') %>% 
  mutate(brand = str_replace_all(brand, ' ', '_'))
  # replaces the " " with '_' enabling easy separation and manipulation of variables.

# Joins the 2 tables into a single complete dataframe using UPC as the key.
wip_cereal_data <- wip_product_data %>% 
  left_join(base_sales_data, 'UPC') 

vtable(wip_cereal_data)

# The following converts variables from class char/numeric to factor as required.
# Base data sets are base set from which additional cleaning, tidying, EAD will be conducted from. 
base_cereal_data <- wip_cereal_data %>% 
  mutate(across(c(maker, brand, package, promo, ad), as.factor))

vtable(base_cereal_data)

write_csv(base_cereal_data, 'base_cereal_data.csv')
```

# Base EDA Step 1: Uni-variate non-graphical EDA

# Base EDA Step 2: Uni-variate graphical EDA

# Based EDA Step 3: Multi-variate non-grapgical EDA

# Base EDA Step 4: Multi-variate graphical EDA

# Save graphs for use in other documents


